---
title: "ggplot Script"
---
```{r load, results='hide', message=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(wesanderson)
library(reshape2)
```

```{r data, include=FALSE}
#import stimuli features file
stim1 <- read.csv(file = "StimuliProperties_Experiment1.csv", head = TRUE, sep = ",")
summary(stim1);
#create variable condition
stim1$condition <- 'condition';
stim1$condition[stim1$primingMorpheme == 'suffix' & stim1$relatedness == 'rel'] <- 'suffix_priming'
stim1$condition[stim1$primingMorpheme == 'suffix' & stim1$relatedness == 'cntr'] <- 'suffix_control'
stim1$condition[stim1$primingMorpheme == 'stem' & stim1$relatedness == 'rel'] <- 'stem_priming'
stim1$condition[stim1$primingMorpheme == 'stem' & stim1$relatedness == 'cntr'] <- 'stem_control'
stim1$condition <- factor(stim1$condition);
stim1$primeFreq <-as.double(stim1$primeFreq);
stim1$primeStemLength <- stim1$PrimeLength-stim1$primeSuffixLength;
stim1$targetStemLength <- stim1$targetLength-stim1$targetSuffixLength;

tstem_mean <- aggregate(targetStemLength ~ condition, FUN=mean, data=stim1);
tstem_sd <- aggregate(targetStemLength~ condition, FUN=sd, data = stim1);
prstem_mean <- aggregate(primeStemLength ~ condition, FUN=mean, data=stim1);
prstem_sd <- aggregate(primeStemLength~ condition, FUN=sd, data = stim1);
tsuffix_mean <- aggregate(targetSuffixLength ~ condition, FUN=mean, data=stim1);
tsuffix_sd <- aggregate(targetSuffixLength~ condition, FUN=sd, data = stim1);
prsuffix_mean <- aggregate(primeSuffixLength ~ condition, FUN=mean, data=stim1);
prsuffix_sd <- aggregate(primeSuffixLength~ condition, FUN=sd, data = stim1);
tfq_mean <- aggregate(log10(targetFreq) ~ condition, FUN=mean, data=stim1);
tfq_sd <- aggregate(log10(targetFreq) ~ condition, FUN=sd, data = stim1);
prfq_mean <- aggregate(log10(primeFreq) ~ condition, FUN=mean, data=stim1);
prfq_sd <- aggregate(log10(primeFreq) ~ condition, FUN=sd, data = stim1);
orth_mean <- aggregate(orthOverlap ~ condition, FUN=mean, data=stim1);
orth_sd <- aggregate(orthOverlap~ condition, FUN=sd, data = stim1)

read.csv(file = "Data_Experiment1.csv", head = TRUE, sep = ",") -> data1;
summary(data1);
#create a variable for target word
data1$target <- data1$TrialID-100;
data1$target[data1$TrialID>199 & data1$TrialID<299] <- data1$target[data1$TrialID>199 & data1$TrialID<299]-100;
data1$target[data1$TrialID>299 & data1$TrialID<399] <- data1$target[data1$TrialID>299 & data1$TrialID<399]-200;
data1$target[data1$TrialID>399 & data1$TrialID<499] <- data1$target[data1$TrialID>399 & data1$TrialID<499]-300;

#factor sbj and items
data1$target <- factor(data1$target);
data1$SubjectID <- factor(data1$SubjectID);

#create new variables "cond": suffix/stem and "relation": control/prime
data1$cond <- 'stem';
data1$cond[data1$Condition == 'suffix_priming'] <- 'suffix';
data1$cond[data1$Condition == 'suffix_control'] <- 'suffix';
data1$cond[data1$Condition == 'nonword'] <- NA;
data1$relation<- 'control';
data1$relation[data1$Condition == 'suffix_priming'] <- 'prime';
data1$relation[data1$Condition == 'stem_priming'] <- 'prime';
data1$relation[data1$Condition == 'nonword'] <- NA;
# convert the new variables to factors
data1$relation <- factor(data1$relation); 
data1$cond <- factor(data1$cond); 

## run the diagnostics, to check for outliers in subjects and items based on RT and accuracy
data1$lexicality = "word"
data1$lexicality[data1$Condition == 'nonword'] <- 'nonword'
data1$lexicality = factor(data1$lexicality)

data1Acc <- subset(data1, Condition != 'nonword' & SubjectID != '13' & target != '39' & RT>0);
summary(data1Acc);
data1Rt <- subset(data1Acc, Accuracy == 1); 
summary(data1Rt);

#convert s to ms on the RT
data1Rt$RTms <- 1000*data1Rt$RT;

#descriptive stats
data_mean <- aggregate(RTms ~ Condition, FUN=mean, data=data1Rt); 
data_sd <- aggregate(RTms ~ Condition, FUN=sd, data = data1Rt);
#checking medians too, RTs are quite long:
aggregate(RTms ~ Condition, FUN=median, data=data1Rt); 
aggregate(RTms ~ Condition, FUN=IQR, data=data1Rt); 
#descriptive stats
acc_mean <- aggregate(Accuracy ~ Condition, FUN=mean, data=data1Acc);
acc_sd <- aggregate(Accuracy ~ Condition, FUN=sd, data=data1Acc)
```



### Stimuli descriptive statistics

#### Graph 1 : Target Words

##### Binding tables

```{r binding1}
target_sd <- left_join(tstem_sd, tsuffix_sd, by = c('condition'))
target_sd <- left_join(target_sd, tfq_sd, by = c('condition'))
target_sd <- melt(target_sd, id=c("condition"))

target_word <- left_join(tstem_mean, tsuffix_mean, by = c('condition'))
target_word <- left_join(target_word, tfq_mean, by = c('condition'))
target_word <- melt(target_word, id=c("condition"))

target_word <- left_join(target_word, target_sd, by = c('condition', 'variable'))
```

##### Cleaning up

Since target words were the same regardless of the condition, I renamed one condition and removed the others.
```{r cleaning1}
target_word<- target_word %>%
  filter(condition == "stem_control")

target_word <- target_word %>%
  rename(SD = value.y)
target_word <- target_word %>%
  rename(value = value.x)

target_word[1] = 'target words'
```

##### Plotting

```{r ggplot1}
plot1 <- ggplot(data = target_word, mapping = aes(x = variable, y = value, fill = variable)) +
  geom_col()+
  geom_errorbar(aes(ymin = value - SD, ymax = value + SD), width = .2) + 
  theme_minimal()+
  scale_x_discrete(labels = c("Stem length",
                              "Suffix length",
                              "Target frequency"),
                         name = NULL) +
  scale_y_continuous(name = NULL,
                    expand = c(0, 0),
                    limits = c(0, 6),
                    breaks = seq(0, 6, 0.5)) +
  scale_fill_manual(values = wes_palette("Moonrise3", n= 3),
                    guide = "none")+
  ggtitle(label = "Average attributes for target words") +
  theme(plot.title = element_text(size = "11",
                                  hjust = "0.5"))

print(plot1)
```

```{r save1, message= FALSE, results='hide'}
ggsave(filename = "targetwords.png", plot = plot1)
```

#### Graph 2 : Orthographic overlap

##### Binding and cleaning

```{r binding2}
orth_sd <- melt(orth_sd, id=c("condition"))
orth_sd <- orth_sd %>%
  rename(SD = value)
orth_mean <- melt(orth_mean, id=c("condition"))
orth <- left_join(orth_mean, orth_sd, by =c("condition", "variable"))

orth <- mutate(orth, relation = ifelse(orth$condition == "stem_control",
                                       "Unrelated prime",
                                       ifelse(orth$condition == "suffix_control",
                                              "Unrelated prime", "Related prime")))

orth <- mutate(orth, cond = ifelse(orth$condition == "stem_control",
                                       "stem",
                                       ifelse(orth$condition == "stem_priming",
                                              "stem", "suffix")))
```

##### Plotting

```{r plot2}
plot2 <- ggplot(data = orth, mapping = aes(x = cond, y = value, fill = cond))+
  geom_col()+
  geom_errorbar(aes(ymin = value - SD, ymax = value + SD), width = .2)+
  facet_wrap(vars(relation), strip.position = "bottom")+
  theme_minimal()+
  scale_x_discrete(labels = NULL,
                   name = NULL)+
  scale_y_continuous(name = NULL)+
  scale_fill_manual(values = wes_palette
                    (name = "GrandBudapest2", n= 2),
                    name = "Condition",
                    labels = c("Stem",
                               "Suffix"))+
  ggtitle(label = "Orthographic overlap between target and prime words")+
  theme(plot.title = element_text(size = "11",
                                  hjust = "0.5"))

print(plot2)
```

```{r save2, message= FALSE, results='hide'}
ggsave(filename = "orthoverlap.png", plot = plot2)
```

#### Graph 3 : Prime frequency

##### Binding and cleaning

```{r binding3}
prfq_sd <- melt(prfq_sd, id=c("condition"))
prfq_sd <- prfq_sd %>%
  rename(SD = value)
prfq_mean <- melt(prfq_mean, id=c("condition"))
prfq <- left_join(prfq_mean, prfq_sd, by =c("condition", "variable"))

prfq <- mutate(prfq, relation = ifelse(prfq$condition == "stem_control",
                                       "Unrelated prime",
                                       ifelse(prfq$condition == "suffix_control",
                                              "Unrelated prime", "Related prime")))

prfq <- mutate(prfq, cond = ifelse(prfq$condition == "stem_control",
                                       "Stem",
                                       ifelse(prfq$condition == "stem_priming",
                                              "Stem", "Suffix")))
```

##### Plotting

```{r plot3}
plot3 <- ggplot(data = prfq, mapping = aes(x = relation, y = value, fill = relation))+
  geom_col()+
  geom_errorbar(aes(ymin = value - SD, ymax = value + SD), width = .2)+
  facet_wrap(vars(cond), strip.position = "bottom")+
  theme_minimal()+
  scale_x_discrete(labels = NULL,
                   name = NULL)+
  scale_y_continuous(name = NULL,
                     breaks = seq(0, 3, 0.5))+
  scale_fill_manual(values = wes_palette
                    (name = "Royal2", n= 2),
                    name = "Relation")+
  ggtitle(label = "Word frequency for prime words")+
  theme(plot.title = element_text(size = "11",
                                  hjust = "0.5"))

print(plot3)
```
```{r save3, message= FALSE, results='hide'}
ggsave(filename = "primefq.png", plot = plot3)
```

#### Graph 4 : Stem length

##### Binding and cleaning

```{r binding4}
prstem_sd <- melt(prstem_sd, id=c("condition"))
prstem_sd <- prstem_sd %>%
  rename(SD = value)
prstem_mean <- melt(prstem_mean, id=c("condition"))
prstem <- left_join(prstem_mean, prstem_sd, by =c("condition", "variable"))

prstem <- mutate(prstem, relation = ifelse(prstem$condition == "stem_control",
                                       "Unrelated prime",
                                       ifelse(prstem$condition == "suffix_control",
                                              "Unrelated prime", "Related prime")))

prstem <- mutate(prstem, cond = ifelse(prstem$condition == "stem_control",
                                       "Stem",
                                       ifelse(prstem$condition == "stem_priming",
                                              "Stem", "Suffix")))
```

##### Plotting

```{r plot4}
plot4 <- ggplot(data = prstem, mapping = aes(x = cond, y = value, fill = cond))+
  geom_col()+
  geom_errorbar(aes(ymin = value - SD, ymax = value + SD), width = .2)+
  facet_wrap(vars(relation), strip.position = "bottom")+
  theme_minimal()+
  scale_x_discrete(labels = NULL,
                   name = NULL)+
  scale_y_continuous(name = NULL,
                    expand = c(0, 0),
                    limits = c(0, 6),
                    breaks = seq(0, 6, 0.5))+
  scale_fill_manual(values = wes_palette
                    (name = "GrandBudapest2", n= 2),
                    name = "Condition")+
  ggtitle(label = "Stem length across condition")+
  theme(plot.title = element_text(size = "11",
                                  hjust = "0.5"))

print(plot4)
```

```{r save4, message= FALSE, results='hide'}
ggsave(filename = "stemlength.png", plot = plot4)
```



#### Graph 5 : Suffix length

##### Binding and cleaning

```{r binding5}
prsuffix_sd <- melt(prsuffix_sd, id=c("condition"))
prsuffix_sd <- prsuffix_sd %>%
  rename(SD = value)
prsuffix_mean <- melt(prsuffix_mean, id=c("condition"))
prsuffix <- left_join(prsuffix_mean, prsuffix_sd, by =c("condition", "variable"))

prsuffix<- mutate(prsuffix, relation = ifelse(prsuffix$condition == "stem_control",
                                       "Unrelated prime",
                                       ifelse(prsuffix$condition == "suffix_control",
                                              "Unrelated prime", "Related prime")))

prsuffix <- mutate(prsuffix, cond = ifelse(prsuffix$condition == "stem_control",
                                       "Stem",
                                       ifelse(prsuffix$condition == "stem_priming",
                                              "Stem", "Suffix")))
```

##### Plotting

```{r plot5}
plot5 <- ggplot(data = prsuffix, mapping = aes(x = cond, y = value, fill = cond))+
  geom_col()+
  geom_errorbar(aes(ymin = value - SD, ymax = value + SD), width = .2)+
  facet_wrap(vars(relation), strip.position = "bottom")+
  theme_minimal()+
  scale_x_discrete(labels = NULL,
                   name = NULL)+
  scale_y_continuous(name = NULL,
                    expand = c(0, 0),
                    limits = c(0, 6),
                    breaks = seq(0, 6, 0.5))+
  scale_fill_manual(values = wes_palette
                    (name = "GrandBudapest2", n= 2),
                    name = "Condition")+
  ggtitle(label = "Suffix length across condition")+
  theme(plot.title = element_text(size = "11",
                                  hjust = "0.5"))

print(plot5)
```

```{r save5, message= FALSE, results='hide'}
ggsave(filename = "suffixlength.png", plot = plot5)
```


### Experimental data descriptive statistics

#### Graph 1 : RT

##### Binding tables

```{r binding6}
data_sd <- melt(data_sd, id=c("Condition"))
data_sd <- data_sd %>%
  rename(SD = value)
data_mean <- melt(data_mean, id=c("Condition"))
dataplot <- left_join(data_mean, data_sd, by =c("Condition", "variable"))
dataplot <- dataplot %>%
  rename(rt = value)

dataplot<- mutate(dataplot, relation = ifelse(dataplot$Condition == "stem_control",
                                       "Unrelated prime",
                                       ifelse(dataplot$Condition == "suffix_control",
                                              "Unrelated prime", "Related prime")))

dataplot <- mutate(dataplot, cond = ifelse(dataplot$Condition == "stem_control",
                                       "Stem",
                                       ifelse(dataplot$Condition == "stem_priming",
                                              "Stem", "Suffix")))
```

##### Plotting

```{r plot6}
plot6 <- ggplot(data = dataplot, mapping = aes())+
  geom_pointrange(aes(x=relation,
                      y=rt, 
                      ymin=(rt-SD), 
                      ymax=(rt + SD),
                      colour = relation))+
  facet_wrap(vars(cond), strip.position = "bottom")+
  theme_minimal()+
  scale_x_discrete(labels = NULL,
                   name = NULL)+
  scale_y_continuous(name = "Response time (ms)",
                     breaks = seq(0, 1000, 50))+
  scale_colour_manual(values = wes_palette
                    (name = "Royal2", n= 2),
                    name = "Relation")+
  ggtitle(label = "Average response time across condition")+
  theme(plot.title = element_text(size = "11",
                                  hjust = "0.5"))

print(plot6)
```

```{r save6, message= FALSE, results='hide'}
ggsave(filename = "dataplot.png", plot = plot6)
```


#### Graph 2 : Accuracy

##### Binding tables

```{r binding7}
acc_sd <- melt(acc_sd, id=c("Condition"))
acc_sd <- acc_sd %>%
  rename(SD = value)
acc_mean <- melt(acc_mean, id=c("Condition"))
acc <- left_join(acc_mean, acc_sd, by =c("Condition", "variable"))
acc <- acc %>%
  rename(Accuracy = value)

acc <- mutate(acc, relation = ifelse(acc$Condition == "stem_control",
                                       "Unrelated prime",
                                       ifelse(acc$Condition == "suffix_control",
                                              "Unrelated prime", "Related prime")))

acc <- mutate(acc, cond = ifelse(acc$Condition == "stem_control",
                                       "Stem",
                                       ifelse(acc$Condition == "stem_priming",
                                              "Stem", "Suffix")))

acc <- mutate(acc, percent = acc$Accuracy*100)

```

##### Plotting

```{r plot7}
plot7 <- ggplot(data = acc, mapping = aes())+
  geom_pointrange(aes(x=Condition,
                      y=percent, 
                      ymin=(percent - SD), 
                      ymax=(percent + SD),
                      colour = Condition))+
  theme_minimal()+
  scale_x_discrete(labels = NULL,
                   name = NULL)+
  scale_y_continuous(name = "Response accuracy %",
                     expand = c(0, 0),
                     limits = c(80, 100),
                     breaks = seq(80, 100, 5))+
  scale_colour_manual(values = wes_palette
                    (name = "Chevalier1", n= 4),
                    name = "Experiment condition",
                    labels = c("Stem unrelated prime",
                               "Stem related prime",
                               "Suffix unrelated prime",
                               "Suffix related prime"))+
  ggtitle(label = "Response accuracy across condition")+
  theme(plot.title = element_text(size = "11",
                                  hjust = "0.5"))

print(plot7)
```

```{r save7, message= FALSE, results='hide'}
ggsave(filename = "accuracy.png", plot = plot7)
```

<style>

h1.title {
  font-size: 32px;
  color: slategrey;
}

h3 {
  font-size: 25px;
  color: grey;
}

h4 {
  font-size: 20px;
  color: darkgrey;
}

p {color: royalblue}

h5 {
    font-size: 15px;
    color: darkslategrey;
}
</style>